CES analysis for Paul Brockway rebound paper

### 2015-08-05 Articles/Brockway_Rebound_2015/Brockway_2015.Rnw
Looking at US and CN results in detail.

@@

@@

@@

The following provides the results in tabular form.

<<resultsTable>>=
fittedParamsTable <- lapply(countries, function(coun){
  cbind(coun, naturalCoef(models[[coun]]))
}) %>% do.call(rbind, .)
@

<<print_resultsTable, results="asis">>=
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE, 
        booktabs=TRUE)
}
myXTable(fittedParamsTable,
         caption="Tabular results.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@

@@

>>>>

First, look a the UK results.

@@

The fitted model for the UK is not near any boundaries
to the economically-meaningful solution space.

Next, look at the US results.

@@

These results for the US deserve additional scrutiny.
The winning US model looks like a boundary model, 
with $\delta = 1$ and $\rho_1 = -1$.
Indeed, this appears to be Boundary Model~7 in \ref{sec:boundary_models}.
Note that $\rho_1 = -1$ corresponds to $\sigma_1 = \infty$.
However, the winning model results report a value for $\sigma$, 
which would be unknowable in Boundary Model~7.

Upon closer examination, we find the following.

<<InvestigateUS, echo=TRUE>>=
# Establish some variable names to make code readable
usdata <- filter(historicalData, Country=="US")
usmodel <- models[["US"]]
# Get the boundary model that is near
bmod7 <- (attr(usmodel, "model.attempts"))$`07:*N1-`
# Did Boundary Model 7 converge?
bmod7$converged
# The fitted parameters in bmod7 are as expected,
# i.e., we get NA for both rho and sigma, 
# and we have delta = 1.
naturalCoef(bmod7)
# Calculate the sse for bmod7
sum(resid(bmod7)^2)
# Compare to sse for fitted model
sum(resid(usmodel)^2)
# So, the sse for the winning model is, indeed, 
# less than the sse for the boundary model.
# To investigate why that is happening,
# let's first look at the coefficient for delta
# in the fitted model.
naturalCoef(usmodel)$delta
# Comes back as 1.
# But machine precision is an issue here. 
# Try this:
naturalCoef(usmodel)$delta - 1
# to find that delta is not EXACTLY 1.00000.
# And, this makes a big difference.
# To investigate further, Randy made a function
# that calculates the ln(y) given historical data and 
# some model coefficients.
logces <- function(coefs, data=usdata) {
  coefs <- as.list(coefs)
  with(coefs,
       log(gamma) + 
         lambda * data$iYear + 
         (-1/rho) * (
           log( 
             delta * ( delta_1 * data$iK^(-rho_1) + (1-delta_1) * data$iL^(-rho_1))^(rho/rho_1) +
               (1-delta) * data$iU^(-rho))
         )
  )
}
# Now, calculate sse in many different ways
# First, the boundary model (repeated from above).
sum(resid(bmod7)^2)
# Next from the winning model (repeated from above).
sum(resid(usmodel)^2)
# Using the logces function
sum((log(usdata$iY) - logces(coef(usmodel)))^2)
# Also try with the boundary model.
# First, need to make an array of coefficients that
# matches the full model
bmod7coeffs <- coef(bmod7)
bmod7coeffs[["gamma"]] <- exp(coef(bmod7)[["logscale"]])
bmod7coeffs[["rho"]] <- 2 # value doesn't matter, because delta = 1.
sum((log(usdata$iY) - logces(bmod7coeffs))^2)
# This result matches the sse value from Boundary Model 7.
# The above tests demonstrate that the logces function is correct.
# Let's understand the sensitivity of sse to 
# the fitted coefficients.
# To do so, we'll calculate sse using rounded values 
# of the model coefficients.
# Try rounding to the 16th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 16)))^2)
# The above works as expected. We
# reproduce the sse value from the model itself.
# Try rounding to the 15th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 15)))^2)
# Same sse value as 16th decimal place.
# Again, works as expected.
# Now, try rounding to the 14th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 14)))^2)
# The sse value is slightly different!
# 13th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 13)))^2)
# 12th decimal place
sum((log(usdata$iY) - logces(round(coef(usmodel), 12)))^2)
# Suddenly, we get a big jump in sse value.
# Note that the value of delta above
coef(usmodel)[["delta"]] - 1
# is on the order of 10^-13,
# right at the point we begin to see 
# significant differences in sse.
@

At this point, Randy and I think that 
our code is correct and we are seeing reality.
We showed that $sse$ is sensitive to very small variations
(in the 13$^{\mathrm{th}}$ decimal place!)
in the values of the fitted coefficients.
Moving slightly away from the boundary
appears to improve the $sse$ enough
so that the full fitted model is superior to 
Boundary Model~7.
And, our software was correct to report
a model that contained all the parameters.

This hypersensitivity to fitted coefficient values
may come about because the value for $\rho$ is so large.
Regardless of the reason, the hypersensitivity
is unnerving (to me). 
And, it provides additional encouragement to
determine the statistical precision of the fitted values.

Finally, look at the CN results.

@@

Similar to the US results above, we can
investigate the CN results in more detail.
It looks like we are on a boundary:
$\delta_1 = 1$ is Boundary Model~11
in \ref{sec:boundary_models}.
But, in fact, we are not.

<<InvestigateCN, echo=TRUE>>=
cndata <- filter(historicalData, Country=="CN")
cnmodel <- models[["CN"]]
naturalCoef(cnmodel)$delta_1
naturalCoef(cnmodel)$delta_1 - 1
# Similar to the US situation, we are close to,
# but not exactly on, the boundary.
# Get the boundary model that is nearby.
bmod11 <- (attr(cnmodel, "model.attempts"))$`11:1-**`
# Calcualte sse for the fitted model.
sum(resid(cnmodel)^2)
# Calculate sse for the nearby boundary model.
sum(resid(bmod11)^2)
@

Again, I think we are looking at reality here.
The full fitted model is better than the nearby boundary model.

I think this all means that we can use the 
UK, US, and CN
results straight away
for rebound calculations.

Tables~\ref{tab:print_resultsTable1} and~\ref{tab:print_resultsTable2}
present the results in readable format.

<<resultsTables>>=
fittedParamsTable1 <- lapply(countries, function(coun){
  cbind(coun, select(naturalCoef(models[[coun]]), scale, lambda, delta, delta_1, rho, rho_1))
}) %>% do.call(rbind, .)

fittedParamsTable2 <- lapply(countries, function(coun){
  cbind(coun, select(naturalCoef(models[[coun]]), alpha_1, alpha_2, alpha_3, sigma_1, sigma))
}) %>% do.call(rbind, .)

myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE, 
        booktabs=TRUE)
}
@

<<print_resultsTable1, results="asis">>=
myXTable(fittedParamsTable1,
         caption="Tabular results (1).",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@

<<print_resultsTable2, results="asis">>=
myXTable(fittedParamsTable2,
         caption="Tabular results (2).",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary models} 
\label{sec:boundary_models}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table} \caption{CES boundary models (rows~1--20) and full model 
(bottom row).} 
\label{tab:CES_boundary_models} 
  \begin{center}
    \begin{tabular}{r c c c c l} 
      \toprule
      & $\delta_1$  & $\sigma_1$ & $\delta$   & $\sigma$ & Boundary model \\
      \midrule
      1. & 1           & ---        & 1          & ---      & $y = \theta A x_1$ 
      \\
      2. & 0           & ---        & 1          & ---      & $y = \theta A x_2$ 
      \\
      3. & ---         & ---        & 0          & ---      & $y = \theta A x_3$ 
      \\
      \midrule
      4. & ---         & 0          & 1          & ---      & $y = \theta A \min 
      (x_1, \, x_2)$ \\
      5. & 1           & ---        & ---        & 0        & $y = \theta A \min 
      (x_1, \, x_3)$ \\
      6. & 0           & ---        & ---        & 0        & $y = \theta A \min 
      (x_2, \, x_3)$ \\
%
      7. &             & $\infty$   & 1          & ---        
            & $y = \theta A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$ \\
%
      8. & 1           & ---        &            & $\infty$        
            & $y = \theta A \left[ \delta x_1 + (1 - \delta) x_3 \right]$ \\
%
      9. & 0           & ---        &            & $\infty$        
            & $y = \theta A \left[ \delta x_2 + (1 - \delta) x_3 \right]$ \\
%
     10. &             &            & 1          & ---
            & $y = \theta A \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/
                    \rho_1}$ \\
%
     11. & 1           & ---        &         & 
            & $y = \theta A \left[ \delta x_1^{-\rho} 
                    + \left(1 - \delta \right) x_3^{-\rho} \right] ^ {-1/\rho}$ 
                    \\ 
%
     12. & 0           & ---        &            & 
            & $y = \theta A \left[ \delta x_2^{-\rho} + \left(1 - \delta \right) 
            x_3^{-\rho} \right]^{-1/\rho}$ \\
      \midrule
     13. & ---         & 0          & ---        & 0        & $y = \theta A \min 
     (x_1, \, x_2, \, x_3)$ \\
%
     14. & ---         & 0          &            & $\infty$        
            & $y = \theta A \left[ \delta \min (x_1, \, x_2) + \left( 1 - \delta 
            \right) x_3  \right]$ \\
%
     15. &             & $\infty$   & ---        & 0
            & $y = \theta A \min \left[ \delta_1 x_1 + \left(1 - \delta_1 
            \right) x_2, \, x_3  \right]$ \\
%
     16. &             & $\infty$   &            & $\infty$        
            & $y = \theta A \left\{ \delta \left[ \delta_1 x_1 + \left( 1 - 
            \delta_1 \right) x_2 \right] 
                    + \left( 1 - \delta \right) x_3  \right\}$ \\
%
     17. & ---         & 0          &            & 
            & $y = \theta A \left\{ \delta \left[ \min \left( x_1, \, x_2 
            \right) \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ 
                    {-1/\rho}$ \\ 
%
     18. &            &            & ---     & 0
            & $y = \theta A \min \left\{ \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/
                    \rho_1}, \, x_3 \right\}$ \\ 
%
     19. &             &            &            & $\infty$
            & $y = \theta A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/
                    \rho_1}  
                    + \left( 1 - \delta \right) x_3  \right\}$ \\ 
%
     20. &             & $\infty$   &            &
            & $y = \theta A \left\{ \delta \left[ \delta_1 x_1 
                    + \left( 1 - \delta_1 \right) x_2 \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ 
                    {-1/\rho}$ \\
\midrule
         &            &            &         & 
            & $y = \theta A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{\rho/
                    \rho_1}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ 
                    {-1/\rho}$ \\ 
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}
